### 2014-05-08 paper/appendix.tex

\chapter{Appendix}
\newpage

\begin{longtabu} to \textwidth { X[1,l] X[6,l] X[4,l] X[4,l]}
\toprule
Pin&Atmel Pin Name&User CPU&Brain CPU\\
\midrule
1&PE6 (INT.6/AIN0)&Digital pin 7&NRF24L01 IRQ\\
2&UVcc&5V&5V\\
3&D-&USB-&USB2-\\
4&D+&USB+&USB2+\\
5&UGnd&GND&GND\\
6&UCap&UCAP&UCAP2\\
7&VUSB&5V&5V\\
8&(SS/PCINT0) PB0&RXLED&RXLED2\\
9&(PCINT1/SCLK) PB1&SCK&SCK2\\
10&(PDI/PCINT2/MOSI) PB2&MOSI&MOSI2\\
11&(PDO/PCINT3/MISO) PB3&MISO&MISO2\\
12&(PCINT7/OCA0/OC1C/\#RTS) PB7&Digital pin 11 (PWM)&Servo Gripper Rotation\\
13&RESET&RESET&RESET2\\
14&Vcc&5V&5V\\
15&GND&GND&GND\\
16&XTAL2&XTAL2&XTAL22\\
17&XTAL1&XTAL1&XTAL12\\
18&(OC0B/SCL/INT0) PD0&SCL&SCL\\
19&(SDA/INT1) PD1&SDA&SDA\\
20&(RX D1/AIN1/INT2) PD2&Digital pin 0 (RX)&RX2\\
21&(TXD1/INT3) PD3&Digital pin 1 (TX)&TX2\\
22&(XCK1/\#CTS) PD5&TXLED&TXLED2\\
23&GND1&GND&GND\\
24&AVCC&AVCC&AVCC\\
25&(ICP1/ADC8) PD4&Digital pin 4&IR\\
26&(T1/\#OC4D/ADC9) PD6&Digital pin 12&Shift Register RCK\\
27&(T0/OC4D/ADC10) PD7&Digital Pin 6 (PWM)&Servo Gripper Grip\\
28&(ADC11/PCINT4) PB4&Digital pin 8&SD Card CS\\
29&(PCINT5/OC1A/\#OC4B/ADC12) PB5&Digital Pin 9 (PWM)&NRF24L01 CE\\
30&(PCINT6/OC1B/OC4B/ADC13) PB6&Digital Pin 10 (PWM)&NRF24L01 CSN\\
31&(OC3A/\#0C4A) PC6&Digital Pin 5 (PWM)&Servo Ultrasonic Rotation\\
32&(ICP3/CLK0/)C4A) PC7&Digital Pin 13 (PWM) + LED&LED\\
33&(\#HWB) PE2&HWB&HWB\\
34&Vcc1&5V&5V\\
35&GND2&GND&GND\\
36&(ADC7/TDI) PF7&Analog In 0&Ultrasonic Trigger\\
37&(ADC6/TDO) PF6&Analog In 1&Ultrasonic Echo\\
38&(ADC5/TMS) PF5&Analog In 2&TFT DC\\
39&(ADC4/TCK) PF4&Analog In 3&TFT RST\\
40&(ADC1) PF1&Analog In 4&TFT CS\\
41&(ADC0) PF0&Analog In 5&MUX Out\\
42&AREF&AREF&AREF\\
43&GND3&GND&GND\\
44&AVCC1&AVCC&AVCC\\
\bottomrule
\caption{CPU pin mapping for constructed robot}
\end{longtabu}

\newpage

\begin{longtabu} to \textwidth { X[3,l] X[1,r] X[1,r] X[1,r]}
\toprule
Part & Amount & Cost per part\footnote{Average price when ordered in quantity of 100.} & Sum \\
\midrule
\multicolumn{4}{c}{\textbf{US Distributor}\footnote{Supplied via \url{http://digikey.de}, accessed on 16th of March 2014. Prices are based on quantities of 100 units.}}\\
\\
CAP 18pF 1206 50V & 4 & 0.04\euro & 0.16\euro \\ 
CAP 0.1uF 1206 50V & 8 & 0.04\euro & 0.32\euro \\
CAP 1uF 1206 50V & 5 & 0.05\euro & 0.25\euro \\
CAP 4.7uF 1206 16V & 2 & 0.06\euro & 0.12\euro \\
CAP 10uF 1206 16V & 3 & 0.05\euro & 0.15\euro \\
CAP 47uF 1206 10V & 2 & 0.28\euro & 0.56\euro \\
RES 22 OHM 1206 1\% & 4 &  0.01\euro & 0.04\euro \\
RES 100 OHM 1206 5\% & 5 & 0.01\euro & 0.05\euro \\
RES 470 OHM 1206 1\% & 1 & 0.01\euro & 0.01\euro \\
RES 1K OHM 1206 5\% & 7 & 0.01\euro & 0.07\euro \\
RES 2K OHM 1206 1\% & 1 & 0.01\euro & 0.01\euro \\
RES 3.09K OHM 1206 1\% & 1 & 0.01\euro & 0.01\euro \\
RES 10K OHM 1206 5\% & 8 & 0.01\euro & 0.08\euro \\
B330A & 1 & 0.25\euro & 0.25\euro \\
LED 1206 & 7 & 0.11\euro & 0.77\euro \\
PTC RESET 15V .5A SMD 1206 & 2 & 0.26\euro & 0.52\euro \\
MOSFET N-CH 20V 4.2A SOT-23 & 2 & 0.19\euro & 0.38\euro \\
FERRITE 3A 100 OHM 1206 SMD & 1 & 0.05\euro & 0.05\euro \\
INDUCTOR 3.3UH 1210 SMD & 1 & 0.15\euro & 0.15\euro \\
CRYSTAL 16MHZ 20PF SMD & 2 & 0.25\euro & 0.50\euro \\
ATMEGA32U4 44TQFP & 2 & 2.80\euro & 5.60\euro \\
LDO 3.3V SOT 23 & 1 & 0.33\euro & 0.33\euro \\ 
MCP1650 & 1 & 0.79\euro & 0.79\euro \\
MCP73831 & 1 & 0.33\euro & 0.33\euro \\
74HC595D SO16 & 2 & 0.20\euro & 0.40\euro \\
74HC4051 SO16 & 1 & 0.21\euro & 0.21\euro \\
ULN2803 SO18 & 1 & 0.54\euro & 0.54\euro \\
\\
\multicolumn{4}{c}{\textbf{Chinese Distributor}\footnote{\label{edragon}Supplied via Chinese distributors found on \url{http://alibaba.com} on the 16th of March 2014. The price is based on quantities of one, dollar prices were converted to Euro with 1\euro = 1.3USD }}\\
\\
CONN 02-JST-XH & 1 & 0.02\euro & 0.02\euro \\
CONN 05-JST-XH & 2 & 0.05\euro & 0.10\euro \\
CONN 2.54 FEMALE 1x06 & 2 & 0.07\euro & 0.14\euro \\
CONN 2.54 FEMALE 1x08 & 1 & 0.08\euro & 0.08\euro \\
CONN 2.54 FEMALE 1x10 & 2 & 0.10\euro & 0.20\euro \\
CONN 2.54 MALE 1x03 & 3 & 0.01\euro & 0.03\euro \\
CONN 2.54 MALE 1x04 & 1 & 0.01\euro & 0.01\euro \\
CONN 2.54 MALE 2x03 & 4 & 0.01\euro & 0.04\euro \\
CONN USB MINI B & 2 & 0.04\euro & 0.08\euro \\
TACTILE SWITCH & 2 & 0.02\euro & 0.04\euro \\
TOGGLE SWITCH & 1 & 0.03\euro & 0.03\euro \\
MICRO SERVO & 2 & 2.04\euro & 4.08\euro \\
IR RECV 38KHZ & 1 & 0.10\euro & 0.10\euro \\
STEPPER MOTOR 28-BYJ & 2 & 1.39\euro & 2.78\euro \\
HC-SR04 ULTRASONIC & 1 & 1.38\euro & 1.38\euro \\
1100mAH LIPO BATTERY & 1 & 1.84\euro & 1.84\euro \\
\\
\multicolumn{4}{c}{\textbf{German Distributor}\footnote{Supplied via German distributors or made in Germany.}}\\
\\
LASERCUT PARTS\footnote{Price is based on the raw material (320x230mm) price of 40\euro per square meter of 3mm acrylic. Lasercutting costs are not included. } & 1 & 3.00\euro & 3.00\euro \\
M4 Screw 25mm & 8 & 0.03\euro & 0.24\euro \\
M4 Screw 10mm & 22 & 0.03\euro &  0.66\euro \\
M4 Nut & 30 & 0.01\euro & 0.30\euro \\
PCB\footnote{Price based on calculation at \url{http://multi-circuit-boards.eu} with the PCB dimensions, double-sided with solder stop.} & 1 & 3.66\euro & 3.66\euro \\ 
\hline \\
Total cost& & & 31.28\euro \\
\bottomrule
\caption{Bill of Material for required parts (assuming mass production of 100 units)}
\label{tbl:bom}
\end{longtabu}

\newpage

\begin{longtabu} to \textwidth { X[3,l] X[1,r] X[1,r] X[1,r]}
\toprule
Part & Amount & Cost per part\footnote{Items supplied via Chinese distributors found on \url{http://alibaba.com} on the 16th of March 2014. The price is based on quantities of one, dollar prices were converted to Euro with 1\euro = 1.3USD } & Sum \\
\midrule
DISPLAY SPI 1.8 INCH & 1 & 4.47\euro & 4.47\euro \\
NRF24L01+ & 1 & 0.92\euro & 0.92\euro \\
MPU6050 GYRO + ACCEL & 1 & 2.76\euro & 2.76\euro \\
HMC-5883L & 1 & 1.84\euro & 1.84\euro \\
MICRO SERVO & 2 & 2.04\euro & 2.04\euro \\
CONN 2.54 FEMALE 1x04 & 1 & 0.04\euro & 0.04\euro \\
CONN 2.54 FEMALE 1x05 & 1 & 0.05\euro & 0.05\euro \\
CONN 2.54 FEMALE 1x08 & 2 & 0.05\euro & 0.10\euro \\
CONN 2.54 FEMALE 2x04 & 1 & 0.05\euro & 0.05\euro \\
\hline \\
Total cost& & & 12.27\euro \\
\bottomrule
\caption{Bill of Material for optional parts (assuming mass production of 100 units)}
\label{tbl:bom_opt}
\end{longtabu}
\includepdf[pages={1}]{images/schematic/robot.pdf}
%\includepdf[pages={1}, angle=90]{images/schematic/robot_schematic.pdf}